---
title: "Assignment Scientific Methods"
author: "Filip Lobpreis"
date: "2024-10-23"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(magrittr)
library(tidyverse)
```








We are loading the historical spending data from [github](https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-02-13/historical_spending.csv)
```{r}
hd = readr::read_csv("historical_spending.csv", col_names = TRUE)
```

# Summary of the historical data

This is a summary of our data set. It shows statistical data such as the minimum and the maximum values for each column, their mean and median values. This way we can easily and quickly visualize these types of data for our dataset. 

```{r}
hd %>% summary()
```


# Data visualization

```{r}
hd %>%
  pivot_longer(PercentCelebrating:GiftCards, names_to = 'param', values_to = 'value') %>%
  ggplot() +
  aes(x = Year) +
  aes(y = value) +
  geom_point() +
  stat_smooth(method='lm') +
  scale_x_continuous(breaks = c(2010, 2015, 2020)) +
  facet_wrap(~param, scales='free') +
  labs(title = "Linear regression over the historical measurements from 2010 until 2022")
```
We are comparing all the categories throughout the years to see how the trend
changes. We can see, that people are buying more things while the time
progresses. The only category that is decreasing is the percentage of people
celebrating. This an be justified by the corona pandemic. However, this trend
occurred even before the pandemic started in 2020.


```{r}
hd %>%
  pivot_longer(PercentCelebrating:GiftCards, names_to = 'param', values_to = 'value') %>%
  ggplot() +
  aes(y = value) +
  geom_boxplot() +
  facet_wrap(~param, scales='free') +
  scale_x_continuous(breaks = c()) +
  labs(title = "Linear regression over the historical measurements from 2010 until 2022")
```



# Testing

```{r}
t.test(hd$GiftCards, hd$GreetingCards)
```

Using the T-test we can determine, that the percentage of people that are buying
the gift cards and greeting cards is not the same. The P-value shows a significant
difference in the two means. Even thou the amount of people buying the cards increases
throughout the years, the percentage is not the same. 
















